Cyclone: Fault Tolerance Middleware for NVM Clusters

### 2016-06-26 paper/paper.tex
Updates to paper.

>>>>

\section{Concurrency}
\subsection{Network}
The single leader in RAFT is responsible for sending every log entry to every
replica. Sending log entries one by one or even in batches causes inefficient
use of the network link between the leader replica and followers when the leader
must wait for an acknowledgment from the follower before sending the next batch
of log entries. The solution is to send log entries from the leader as they are
appended to its log without waiting for an acknowledgment. The result however is
that entries can arrive out of order at follower replicas due to an unreliable
network.

We note that RAFT as specified already tolerates out of order arrival of log
entries at follower replicas (a newly elected leader assumed replicas are up to
date when it starts sending out log entries). However follower replicas can only
apply log entries \emph{in order}- dropping log entries that cannot be applied
to extend the log at a follower replica. To bridge this gap, with no changes to
the protocol, we added an in-memory cache of log entries (that we dub the log
cache) at follower replicas to cache log fragments that arrive out of
order. These are then applied opportunistically as soon as all preceding entries
have arrived.

Pseudocode for the log cache is shown in Figure~\ref{fig:logcache}.

\begin{figure}
{ \scriptsize
\begin{verbatim}
on ReceiveEntry(leaderTerm, entry) {
  if (leaderTerm < logcache_term) {
    drop entry // Stale
  }
  else if(leaderTerm > logcache_term){
    drop all entries from logcache
    logcache_term = leaderTerm
  }
  while(entry = canAdd(raftLog, logcache)) {
    AddEntry(raftLog, entry);
    RemoveEntry(logcache, entry)
    Send RAFT response for entry;
  }
}
\end{verbatim}
}
\caption{Logcache}
\label{fig:logcache}
\end{figure}

The log cache is built around the key property of RAFT that there is exactly one
leader in any term and the log at the leader of a term can only be appended to. We
therefore need only support adding entries to the log cache and flush the entire
log cache on detecting a change in leadership due to a term larger than that of
the leader whose entries are cached. Application of log entries follows the
normal RAFT protocol where the last entry in the log must match the desired
previous log index and previous log entry term specified in the new entry to be
appended.

The log cache enables leaders to send batches of log entries as soon as they are
available rather than waiting for an acknowledgment from followers for the
previous batch - enabling better utilization of the network link. We note
however that the leader continues to track the last log entry \emph{known} to be
replicated at any follower (the {\tt match index} in the RAFT protocol
specification) as usual and resends log fragments on timeouts - tolerating
unreliable networks that drop log fragments.

In addition to better utilization of the network link out of the leader replica,
the log cache also enables the leader replica to effectively offload the work of
replication. To do so it simply informs the client originating the log entry of
the list of replicas to which the entry must be sent. The client sends the
entries, collects the responses and responds to the leader replica as soon as a
majority quorum has responded. From the perspective of the RAFT protocol the
client simply acts as an unreliable router. The log cache allows follower
replicas to utilize chunks that might arrive out of order from different
assisting clients. The benefit of such a scheme is that the replication load on
the leader (in terms of network bandwidth) is distributed across the clients
with a larger load (more clients) permitting more offload. As we show in the
evaluation this is an effective technique to keeping tail latency under control
in Cyclone deployments that can tolerate the loss in performance due to the
extra hop but are willing to pay that price for better tail latencies.
